### 2024-10-30 main_widom_morse.tex

\documentclass[12pt,a4paper]{article}

\usepackage[english]{babel}
\usepackage[T2A]{fontenc}
\usepackage[cp1251]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{color}
\usepackage{amsfonts}
\usepackage{wrapfig}
\usepackage{caption}

\usepackage{hyperref}
\usepackage{physics}

\textheight=26cm
\textwidth=17cm
\hoffset=0pt
\voffset=0cm
\oddsidemargin=0cm
\topmargin=-60pt
\footskip=30pt
\renewcommand{\baselinestretch}{1}

\newcommand{\no}{\nonumber}
\newcommand{\non}{\nonumber \\}
\newcommand{\ve}[1]{{\bf #1}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\sli}{\sum\limits}
\newcommand{\ili}{\int\limits}
\newcommand{\lp}{\left (}
\newcommand{\rp}{\right )}
\newcommand{\lb}{\left [}
\newcommand{\rb}{\right ]}
\newcommand{\lbr}{\left \{}
\newcommand{\rbr}{\right \}}
\newcommand{\ld}{\left .}
\newcommand{\rd}{\right .}
\newcommand{\vk}{\vec{k}}
\newcommand{\vl}{\vec{l}}
\newcommand{\cB}{{\cal B}}
\newcommand{\cM}{{\frak M}}
\newcommand{\rhok}{\rho_{\vec{k}}}
\newcommand{\rhomk}{\rho_{-\vec{k}}}
\newcommand{\etak}{\eta_{\vec{k}}}
\newcommand{\etamk}{\eta_{-\vec{k}}}

\begin{document}

\binoppenalty=10000
\relpenalty=10000
\begin{center}
	\renewcommand{\theequation}{\arabic{section}}
	\section*{Supercritical Crossover Lines in the Cell Fluid Model}
	\setcounter{equation}{0}
\end{center}

\subsection*{Abstract}
A cell fluid model with a modified Morse potential is studied. The supercritical states are considered with respect to a possibility to build a separation boundary between liquid-like and gas-like bahaviors. Three different lines are calculated that can be used for this purpose: the locus of the isothermal compressibility maxima, the locus of the thermal expansion coefficient maxima, and the line where the effective chemical potential is zero, $M=0$. By the symmetry of the functionals for the partition functions, the condition $M=0$ in fluids is analogous to the absence of an external field in the Ising model.

% cell fluid model, Morse potential, supercritical region, Widom line

% Продовжується дослідження коміркової моделі плину з модифікованим потенціалом Морзе. Надкритичні стани розглядаються з огляду на можливість побудови межі розділення між рідиноподібними та газоподібними поведінками. Розраховано три різні лінії, які можна використовувати для цієї мети: лінія максимумів ізотермічної стисливості, лінія максимумів коефіцієнта теплового розширення та лінія, де ефективний хімічний потенціал дорівнює нулю, $M=0$. За симетрією функціоналів статистичних сум умова $M=0$ у рідинах аналогічна відсутності зовнішнього поля в моделі Ізінга. {\textit{К\,л\,ю\,ч\,о\,в\,і\,	с\,л\,о\,в\,а:} коміркова модель, потенціал Морзе, надкритична область, лінія Відома.}

\input{./paper_text.tex}

%\subsubsection{Лінія Відома (лінія максимумів стисливості)}

%\begin{figure}[h!]
%	\centering \includegraphics[width=0.5\textwidth]{f5.pdf}
%	\vskip-3mm\caption{Plot of the isothermal compressibility maxima line (the Widom line).	}\label{fig5}
%\end{figure}

%\subsubsection{Лінія розмежування газоподібної і рідиноподібної структур надкритичного плину, виходячи з рівняння стану при $M=0$}
%\begin{figure}[h!]
%	\centering \includegraphics[width=0.5\textwidth]{f6.pdf}
%	\vskip-3mm\caption{Plot of the crossover line between gas-like and liquid-like structures of the supercritical cell fluid based on the equation of state.	}\label{fig6}
%\end{figure}
%\subsubsection{Порівняння обох ліній}

%\bibliographystyle{elsarticle-num}
%\bibliographystyle{plain}
\bibliographystyle{gost2008}
\bibliography{extracted}

\end{document}

\begin{thebibliography}{99}

	%1
	\bibitem{KMLetal14} {\v{Z}}.~Knez, E.~Marko{\v{c}}i{\v{c}}, M.~Leitgeb, M.~Primo{\v{z}}i{\v{c}}, M.~Knez Hrn{\v{c}}i{\v{c}}, M.~{\v{S}}kerget.
	\href{https://doi.org/10.1016/j.energy.2014.07.044}{Industrial applications of supercritical fluids: A review}. \textit{Energy} \textbf{77} 235 (2014) [doi: 10.1016/j.energy.2014.07.044]
	%2
	\bibitem{FW69} M.E.~Fisher, B.~Widom.
	\href{https://doi.org/10.1063/1.1671624}{Decay of Correlations in Linear Systems}. \textit{J. Chem. Phys.} \textbf{50} 3756 (1969) [doi: 10.1063/1.1671624]
	%3
	\bibitem{BGMetal17} T.~Bryk, F.A.~Gorelli, I.~Mryglod, G.~Ruocco, M.~Santoro, T.~Scopigno.
	\href{https://doi.org/10.1021/acs.jpclett.7b02176}{Behavior of Supercritical Fluids across the “Frenkel Line”}. \textit{J. Chem. Phys. Lett.} \textbf{8} 4995 (2017) [doi: 10.1021/acs.jpclett.7b02176]
	%4
	\bibitem{CKS01} X.~Campi, H.~Krivine, N.~Sator.
	\href{https://doi.org/10.1016/S0378-4371(01)00158-3}{Percolation line of self-bound clusters in supercritical fluids}. \textit{Physica A} \textbf{296} 24 (2001) [doi: 10.1016/S0378-4371(01)00158-3] 
	%5
	\bibitem{XKBetal05} L.~Xu, P.~Kumar, S.V.~Buldyrev, S.-H.~Chen, P.H.~Poole, F.~Sciortino, H.E.~Stanley.
	\href{https://doi.org/10.1073/pnas.0507870102}{Relation between the Widom line and the dynamic crossover in systems with a liquid-liquid phase transition}. \textit{PNAS} \textbf{102} 16558 (2005) [doi: 10.1073/pnas.0507870102]
	%6
	\bibitem{MS10} P.~McMillan, H.~Stanley.
	\href{https://doi.org/10.1038/nphys1711}{Going supercritical}. \textit{Nature Phys.} \textbf{6} 479 (2010) [doi: 10.1038/nphys1711]
	%7
	\bibitem{AKM17} S.~Artemenko, P.~Krijgsman, V.~Mazur. 
	\href{https://doi.org/10.1016/j.molliq.2017.03.107}{The Widom line for supercritical fluids}. \textit{J. Mol. Liq.} \textbf{238} 122 (2017) [doi: 10.1016/j.molliq.2017.03.107]
	%8
	\bibitem{Proctor20} J.E.~Proctor.
	\href{https://doi.org/10.1201/9780429491443}{The liquid and supercritical fluid states of matter}. (CRC Press, 2020)  [ISBN: 9780429491443]
	%9
	\bibitem{KD20} M.~Kozlovskii, O.~Dobush.
	\href{https://ujp.bitp.kiev.ua/index.php/ujp/article/view/2019631}{Phase behavior of a cell fluid model with modified Morse potential.} \textit{Ukr. J. Phys.} \textbf{65}, 428 (2020) [doi: 10.15407/ujpe65.5.428]
	%10
	\bibitem{DKRP24} O.A.~Dobush, M.P.~Kozlovskii, R.V.~Romanik, I.V.~Pylyuk.
	\href{https://doi.org/10.48550/arXiv.2409.09786}{Thermodynamic response functions in a cell fluid model}. \textit{Ukr. J. Phys} - in print [doi: 10.48550/arXiv.2409.09786]
	%11
	\bibitem{DKRP24arxiv} O.A.~Dobush, M.P.~Kozlovskii, R.V.~Romanik, I.V.~Pylyuk.
	\href{https://doi.org/10.48550/arXiv.2409.09786}{Thermodynamic response functions in a cell fluid model}. \textit{arXiv preprint} 2409.09786 (2024) [doi: 10.48550/arXiv.2409.09786]
	%12
	\bibitem{Morse29} P.~M.~Morse. 
	\href{https://link.aps.org/doi/10.1103/PhysRev.34.57}{Diatomic molecules according to the wave mechanics. II. Vibrational levels.} \textit{Phys. Rev.} \textbf{34}, 57 (1929) [doi: 10.1103/PhysRev.34.57]
	%13
	\bibitem{SBGetal10} G.G.~Simeoni, T.~Bryk, F.A.~Gorelli, M.~Krisch, G.~Ruocco, M.~Santoro, T.~Scopigno.
	\href{https://doi.org/10.1038/nphys1683}{ The Widom line as the crossover between liquid-like and gas-like behaviour in supercritical fluids}. \textit{Nature Phys.} \textbf{6} 503 (2010) [doi: 10.1038/nphys1683]
	%14
	\bibitem{KozlovskiiDobush2016} M.~Kozlovskii, O.~Dobush.
	\href{https://doi.org/j.molliq.2015.12.018}{Representation of the grand partition function of the cell model: The state equation in the mean-field approximation.} \textit{J. Mol. Liq.} \textbf{215}, 58 (2016) [doi: j.molliq.2015.12.018]
	%15
	\bibitem{KR09} M.P~Kozlovskii, R.V.~Romanik.
	\href{https://doi.org/10.30970/jps.13.4007}{The equation of state of a three-dimensional {Ising}-like system}. \textit{Journal of Physical Studies} \textbf{13} 4007 (2009) [doi: 10.30970/jps.13.4007]
	%16
	\bibitem{Kulinskii10jcp} V.L.~Kulinskii.
	\href{https://doi.org/10.1063/1.3457943}{Global isomorphism between the {Lennard-Jones} fluids and the {Ising} model}. \textit{J. Chem. Phys.} \textbf{133} 034121 (2010) [doi: 10.1063/1.3457943]
	%17
	\bibitem{BK11} L.A.~Bulavin, V.L.~Kulinskii.
	\href{https://doi.org/10.1021/jp201872f}{Unified Picture for the Classical Laws of {B}atschinski and the Rectilinear Diameter for Molecular Fluids}. \textit{J. Phys. Chem. B} \textbf{115} 6061 (2011) [doi: 10.1021/jp201872f]

\end{thebibliography}

%\end{document}